Hierarchical Dispatching and Overriding

### 2017-06-15 Aplas17/paper.tex

% \title{Untangling Unintended Conflicts with Hierarchical Dispatch}
% \title{FHJ: A Formal Model for Resolving Unintended Conflicts in Multiple Inheritance}
\title{Hirarchical Dispatching and Overriding}
%\subtitle{\InterfaceBased Programming for the Masses}

@@

This paper presents a new model \name{} 
(short for \emph{\textbf{F}eatherweight \textbf{H}irarchical \textbf{J}ava}) 
that deals with unintentional
method conflicts. In our new model conflicting methods arising
from unrelated methods can coexist in the same component. Moreover they
can be \emph{independently overridden}. What ensures unambiguity 
is the use of information about the class hierarchy. In turn
hierarchical information is also employed in method dispatching, which 
uses a combination of static and dynamic type information to choose 
the implementation of a method at run-time. We give illustrative 
examples of our language and features and formalize \name{} as a minimal 
Featherweight-Java style calculus.

>>>>

% \title{Untangling Unintended Conflicts with Hierarchical Dispatch}
% \title{FHJ: A Formal Model for Resolving Unintended Conflicts in Multiple Inheritance}
\title{Hierarchical Dispatching and Overriding}
%\subtitle{\InterfaceBased Programming for the Masses}

@@

This paper presents a new model \name{} 
(short for \emph{\textbf{F}eatherweight \textbf{H}ierarchical \textbf{J}ava}) 
that deals with unintentional
method conflicts. In our new model conflicting methods arising
from unrelated methods can coexist in the same component. Moreover they
can be \emph{independently overridden}. What ensures unambiguity 
is the use of information about the class hierarchy. In turn
hierarchical information is also employed in method dispatching, which 
uses a combination of static and dynamic type information to choose 
the implementation of a method at run-time. We give illustrative 
examples of our language and features and formalize \name{} as a minimal 
Featherweight-Java style calculus.

### 2017-06-15 Aplas17/sections/Introduction.tex

A reason why programming languages have resisted to multiple
inheritance in the past is that, as Cook~\cite{Cook1987} puts it, 
multiple inheritance is good but ``\emph{there is no good way to do it}''.
One of the most sensitive and critical issues is perhaps the ambiguity
introduced by multiple inheritance. One case is the famous
\textit{diamond problem}~\cite{Sak89dis,Singh1995}. In the diamond problem inheritance could allow
one feature to be inherited from multiple parent classes that share a
common ancestor, hence
conflicts arise. The variety of strategies for resolving such conflicts
urges the occurrence of different multiple inheritance models,
including traits, mixins, CZ~\cite{malayeri2009cz}, and many others. Existing
languages and research have taken care of this case extensively. Other issues
include how multiple inheritance deals with state, 
which also has been discussed quite extensively~\cite{classless,malayeri2009cz,stroustrup1995}.

>>>>

A reason why programming languages have resisted to multiple
inheritance in the past is that, as Cook~\cite{Cook1987} puts it, 
``\emph{multiple inheritance is good but there is no good way to do it}''.
One of the most sensitive and critical issues is perhaps the ambiguity
introduced by multiple inheritance. One case is the famous
\textit{diamond problem}~\cite{Sak89dis,Singh1995} (also known as ``fork-join inheritance''~\cite{Sak89dis}). In the diamond problem inheritance could allow
one feature to be inherited from multiple parent classes that share a
common ancestor, hence
conflicts arise. The variety of strategies for resolving such conflicts
urges the occurrence of different multiple inheritance models,
including traits, mixins, CZ~\cite{malayeri2009cz}, and many others. Existing
languages and research have taken care of this case extensively. Other issues
include how multiple inheritance deals with state, 
which also has been discussed quite extensively~\cite{classless,malayeri2009cz,stroustrup1995}.

### 2017-06-15 Aplas17/sections/RelatedWork.tex

\subsection{Resolving Unintended Method Conflicts}
There are a few languages that partly realize the problem of
unintended conflicts and provide some support for it.

\noindent {\bf C++ model.}
C++ supports a very flexible inheritance model and allows programmers to choose either static dispatch or dynamic dispatch for method lookup.
It allows unintended conflicts and uses static dispatch to resolve them, as discussed in Section~\ref{sec:overview}. For example, given the following code
\begin{lstlisting}[language=Java]
class A { public: void m() {cout << "MA" << endl;} };
class B { public: void m() {cout << "MB" << endl;} };
class C : public A, public B { 
	void m() {cout << "MC" << endl;}
};
void func(A* a) { a->m(); }
int main() {
	C* c = new C();
	c->B::m();
	func(c); 
	return 0; //Running result: MB MA
}
\end{lstlisting}
The running result is $MB \; MA$, meaning that it uses static dispatch (looks at the static types) for method \lstinline|m()|. 
On calling \texttt{func(c)}, in spite that the dynamic type of \texttt{c} is class \texttt{C}, the method call still dispatches to 
\lstinline|A.m|.
However, we can alter the code a little bit with \textbf{\texttt{virtual}} methods and the result will be totally different:
\begin{lstlisting}[language=c++]
class A { public: virtual void m() {cout << "MA" << endl;} };
class B { public: virtual void m() {cout << "MB" << endl;} };
class C : public A, public B { 
    public: virtual void m() {cout << "MC" << endl;}
};
void func(A* a) { a->m(); }
int main() {
	C* c = new C();
	c->B::m();
	func(c); 
	return 0; //Running result: MB MC
}
\end{lstlisting}
Now the running result will be $MB \; MC$. With virtual methods, dynamic dispatch is used and 
method lookup algorithm will find the most specific method definition of $m$, namely \lstinline|C.m| at this time.
Although C++ support this flexibility, dynamic dispatching on
conflicting methods is problematic.\bruno{why? give example of
  something that goes wrong?}
Furthermore, C++ does not support partial overrides compared to our
model.

>>>>

\subsection{Resolving Unintentional Method Conflicts}
There are a few languages that partly realize the problem of
unintentional conflicts and provide some support for it.

\noindent {\bf C++ model.}
C++ supports a very flexible inheritance model and allows programmers to choose either static dispatch or dynamic dispatch for method lookup.
It allows unintentional conflicts and uses static dispatch to resolve them, as discussed in Section~\ref{sec:overview}. For example, given the following code
\begin{lstlisting}[language=Java]
class A { public: void m() {cout << "MA" << endl;} };
class B { public: void m() {cout << "MB" << endl;} };
class C : public A, public B { 
	void m() {cout << "MC" << endl;}
};
void func(A* a) { a->m(); }
int main() {
	C* c = new C();
	c->B::m();
	func(c); 
	return 0; //Running result: MB MA
}
\end{lstlisting}
The running result is $MB \; MA$, meaning that it uses static dispatch (looks at the static types) for method \lstinline|m()|. 
On calling \texttt{func(c)}, in spite that the dynamic type of \texttt{c} is class \texttt{C}, the method call still dispatches to 
\lstinline|A.m|.
However, we can alter the code a little bit with \textbf{\texttt{virtual}} methods and the result will be totally different:
\begin{lstlisting}[language=c++]
class A { public: virtual void m() {cout << "MA" << endl;} };
class B { public: virtual void m() {cout << "MB" << endl;} };
class C : public A, public B { 
    public: virtual void m() {cout << "MC" << endl;}
};
void func(A* a) { a->m(); }
int main() {
	C* c = new C();
	c->B::m();
	func(c); 
	return 0; //Running result: MB MC
}
\end{lstlisting}
Now the running result will be $MB \; MC$. With virtual methods, dynamic dispatch is used and 
method lookup algorithm will find the most specific method definition of $m$, namely \lstinline|C.m| at this time.
Although C++ support this flexibility, dynamic dispatching on
conflicting methods is problematic.\bruno{why? give example of
  something that goes wrong?}
Furthermore, C++ does not support partial overrides compared to our
model.
